Passenger Prediction in Shared Accounts for Flight Ticket Recommendation

### 2016-06-28 Paper.tex
Add files via upload

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission STATEment
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate-05-2015}

\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

\acmPrice{\$15.00}

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-eSTATE'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
XXX\\
       \affaddr{XXX}\\
       \affaddr{XX}\\
       \email{XXX}
% 2nd. author
\alignauthor
XXX\\
       \affaddr{XXX}\\
       \affaddr{XX}\\
       \email{XXX}
% 3rd. author
\alignauthor
XXX\\
       \affaddr{XXX}\\
       \affaddr{XX}\\
       \email{XXX}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Personalized recommendation is needed for online flight ticket booking service because it is a difficult task for a traveller to select the flight when the number of available flights is large. Traditionally, we can recommend flights for a user based on his historical orders from his account. However, people often book tickets for companions through his account. In this case, the preferences of other travellers should also be considered. Unfortuanately, before placing the order, people will not provider passengers' information. Therefore, we propose a probabilistic method for passenger prediction based on historical behaviors and contextual knowledge. We then experimentally demonstrate its effectiveness on a real dataset. The result shows that our method outperforms conventional methods.
\end{abstract}

\keywords{recommender system, probabilistic topic model, shared account}

\section{Introduction}
Recent years, with the rapid development of online travel agency, there are more and more passengers booking flight tickets through online service providers. Typically, an user may input his location, destination and departure time, then he will get a list of candidate flights. In general, there are dozens of items within a list, all of them are isomorphic. Thus, it takes quite long time for the user to compare and choose the most appropriate one. Therefore, a recommender system is quite necessary to be applied for better user experience. Figure 1 shows a list for all candidate flights departure from Beijing to Hong Kong when input a query. There are total 156 matchable flights in result. The website only provides simple sorting strategies such as departure time, arrival time, price. Thus, it may take quite long time for users to find the appropriate one.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.48]{flight_tickets.eps}
\caption{Search result for flight tickets}
\label{fig:Prob}
\end{figure}\par
Essentially, an item recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that every account represents a single preference model. However, in online flight ticket booking, it is quite common that a user books tickets for the whole family or companions.  Obviously, every passenger has his own preferences. Unfortuanately, before placing the order, people will not provider passengers' information.  Therefore, if we want to improve the accuracy of recommendation, we have to predict who are going to fly.  \par
To predict who is using the account to place orders is regarded as shared account problem. Some approaches have been proposed to extract implicit user information within an account. For example,  in \cite{kabbur:nlmf}\cite{jason:embedding} nonlinear latent factorization is applied to  extract user information. However, these methods only rely on the analysis of relationships between products and users \cite{koen:top-n}\cite{yutaka:modeling}, and context information that can help differentiate users is neglected. Besides, current methods always treat an account as a fixed number of implicit users, and the experiment result can be only evaluated on artificial composite datasets. 
The flight booking service providers can acquire user's personal information only after he has submitted the order. Thus, our goal is to predict distribution of passengers in current session based on historical behavior and contextual knowledge so that we can make more pertinent recommendations. To achieve this target, we adopt the probabilistic author topic model to analyze passengers' behaviors. We regard all orders in an account as corpus, all passengers appearing in the account as authors, orders as documents, content attributes of the items and contextual knowledge as item list. Note that if an order has more than one passenger, they are co-authors of this document.\par
In experiment section, we apply our method to a real flight ticket order dataset.  And the result shows that our method has higher identification accuracy and a better recommendation accuracy.\par
This paper has following contributions:\\
\begin{enumerate}
\item We provide a probabilistic passenger identification method for online flight booking service.
\item We apply the passenger prediction results to flight recommendation.
\item We evaluate our approach on a real flight order dataset, and the result shows that our approach can effectively improve the recommendation accuracy.
\end{enumerate}
\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in Section 2. In Section 3, we define the problem and describe the model. Besides, we also discuss how to do parameter inference. In section 4, we introduce the experiment dataset, evaluation metrics and the experimental results of passenge prediction and flight recommendation. Finally, we conclude the paper in section 5.

\section{Related Work}
\textbf{Content based recommendation.} In online travel service industry such as airline tickets or hotel booking, the item information is usually highly structured, however, this information is quite mutable. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper representations of the attributes of items and the profile of user interests, then it tries to match  the profile of user's interests with attributes of items. In general, a content-based recommender system consists of three components. A content analyzer extracts structured attributes of items through a series of pre-processing steps. It produces the inputs for the system, which always need domain related knowledge. A profile learner analyzes and constructs user's profile of interests by collecting user's historical behaviors. In addition, this component also relies on feedbacks of users. A filtering component evaluates the relevance of items to particular users by matching them with users' profiles. The output for an item can be either binary or continuous, in this stage, some similarity measurements including cosine or euclidean distance can usually be applied.\par

\textbf{Flight Ticket Recommendation.} There is little work focusing on personalized flight ticket recommendation. Many online flight ticket booking agencies provide sorting, filtering strategies to improve user experience, such as sorting by price or takeoff time and filtering by airline or departure airport. Others also provide additional information that can help users, for example, placing the hottest flights at top or highlighting some of them. These methods do make sense in helping users find appropriate flights within less time. However, they can not provide personlized recommendations. In 2004, Lorcan Coyle\cite{lor:flight} proposed a Personal Travel Assistant(PTA). PTA is a method of case based reasoning, which takes user's historical orders as cases and recommends the most similar candidate flight to user's query. In addition, PTA also proposes attribute weight training and case storage method.\par
\textbf{Implicit feedback.} The user's feedback can be divided into implict  or explict type. The explicit feedback can be acquired from users' ratings or text comments. However,  users may not willing to provide explict feedback since it needs additional efforts. Moreover, in many applications, a single numeric rating may not be adequate for describing users feeling to an item. Implicit feedback methods\cite{song:mining} try to infer users' attitudes to an item based on his behaviors, such as clicking ,collecting or bookmarking et al. It has the advantage of being free from the problem of lacking feedback data. Nevertheless, implicit feedback based recommender suffers from problems of lack of negative feedbacks and data noises. In order to provide recommendations when only implict feedbacks are available, some methods are proposed, such as one class collaborative filtering and content based recommender.\par
\textbf{Probabilistic topic models.}  Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data, which are widely used in information retrieval, document annotation, etc. The main step is extracting latent factors from corpus, named topics, which are commonly probabilistic distribution over words. Topic models have three major advantages over other document modeling methods. At first, topic models provide a completely unsupervised approach to extract topics, thus requiring no document labels and no initialization. Unsupervised methods are necessary for modeling large document corpus, besides, in some fields, with the rapid change of content's topics, predefined topic categories may not reflect the development of specific field. Secondly, latent topics are interpretable respectively, so that extracted results are understandable for users. Finally, each document may consist of multiple topics, document can be considered as probabilistic distribution over topics. Many existing models regard documents as mixtures of topics, such as pLSI, LDA. The author topic model is a generative model that extends LDA to include authorship information for document modeling. In our approach, topic model is innovatively applied to indentify users. \par
\textbf{Shared account recommendation.} Some work aims at addressing the challenge of identifying current users who share a single account. A top-N recommendation for shared accounts approach was proposed by Koen et al.\cite{koen:top-n} It is an item-based top-N collaborative filtering recommender system on binary, positive-only feedback. Another method proposed by Santosh et al.\cite{kabbur:nlmf} models users within an account with a much richer representation. It uses a nonlinear matrix factorization methods for predicting the recommendation score. Yutaka et al.\cite{yutaka:modeling} introduced an approach for modeling multiple users' purchase in a single account using an extended pLSA model. All above methods are applied in scenario that users' personal information can't be acquired explicitly.

\section{Passenger Prediction Model}
In this Section, we propose a generic model for predicting the probabilistic distribution of passengers under certain contextual environment. Finally we integrate user prediction into recommendation to get higher accuracy. Figure 2 shows an overview of the recommendation process. At first,a dataset for orders are stored, each account may contains several passengers. We use this data to train user prediction model as well as extract preference at user granularity. Before a recommendation process starts, we calculate user distribution probability applying trained model and contextual knowledge of current session. At the final step, we generate a ranked candidate list as the recommendation result based on user prediction result and user's preference.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of flight recommendation based on passenger prediction}
\label{fig:Prob}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items, it represents the latent feature. And an account is a multinomial distribution over topics, it can represent the preference of an account. Each purchase can be regared a sample that selects a topic z from account a, and takes an item from topic z.\par

When topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite different to be modeled. Conventional items only have static features, means that the content of an item is less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the ticket of different price as one item. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly.\par

Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. Some factors are discrete while others are continuous. For continuous factors, we can partition them into discrete non-overlapping intervals applying . Thus, the number of discrete alternatives of each factor is also very limited. We take orders within each account as a corpus. Our notations are summarized in table 1.\par
\begin{table}[!htbp]
\centering
\caption{Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected Factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

@@

\begin{enumerate}
\item For each passenger $p=1,\dots,|P|$ draw $\Theta_p \sim Dirichlet(\alpha)$
\item For each topic $t=1,\dots,K$ draw $\Phi_t \sim Dirichlet(\beta)$
\item For each order $o=1,\dots,O$
       \begin{enumerate}[fullwidth,itemindent=1em,label=(\alph*)]
       \item given passengers $P$
       \item For each word $i=1,\dots, N_o$
              \begin{enumerate}[fullwidth,itemindent=2em,label=(\roman*)]
              \item draw a passenger $X_{oi} \sim Uniform(P)$
              \item draw a topic $Z_{oi} \sim Discrete(\theta_{X_{oi}})$
              \item draw a word $w_{oi} \sim Discrete(\phi_{Z_{oi}})$
              \end{enumerate}
       \end{enumerate}
\end{enumerate}
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.55]{f1.eps}
\caption{Probability Graph for the author-topic model}
\label{fig:Prob}
\end{figure}\par
Figure 3 shows the graphical model corresponding to this process, where shaded and non-shaded nodes represent observed and unobserved variables respectively. It's a plate notation, the number at the bottom of each box indicates repeated times of operation. Arrows represents the conditional dependency between variables. In the author-topic model, observed variables include words in an order and the passengers of the order.

\subsection{Parameter inference}
As mentioned above, the author-topic model includes two sets of unknown parameters, the $P$ passenger-topic distributions $\theta$, and the $K$ topic-word distributions $\phi$. There are a variety of algorithms that can be applied to estimate the parameters of author-topic models, the Expectation-Maximization algorithm and Gibbs sampling\cite{gregor:esti}. Generic EM algorithms are likely to face local maxima and computational inefficient. In this paper, we utilize Gibbs sampling, it does not explicitly estimate parameters, instead, it evaluates posteriori distribution just on drawn passenger $X$ and topic $Z$. Thus it is simple for Dirichlet priors.\par
We can obtain the probability of every word $\mathbf{w}_o$ generated in each order, conditioned on $\Theta$ and $\Phi$ is:\\
\begin{flalign}
\begin{split}
& P(\mathbf{w}_o | \Theta,\Phi,P) =\prod_{i=1}^{N_o}P(w_{oi}|\Theta,\Phi,\mathbf{p}_o) \\
& =\prod_{i=1}^{N_o}\sum_{p=1}^{|P|}\sum_{t=1}^{K}P(w_{oi}|z_{oi}=t,\Phi)
P(z_{oi}=t|x_{oi}=p,\Theta)P(x_{oi}=p|\mathbf{p}_o)\\
& =\prod_{i=1}^{N_o}\frac{1}{|P|}\sum_{p \in p_o}\sum_{t=1}^{K}\phi_{w_{oi}t}\theta_{tp}
\end{split} &
\end{flalign}

With the above generative model, $P(x_{oi}=p|\mathbf{p}_o)$ is assumed to be a uniform distribution over passenger list $P$. Each topic is drawn independently when conditioned on $\Theta$ and $p_o$, and each word is drawn independently when conditioned on $\Phi$ and $z$. Equation 1 can be applied as the likelihood of all orders in a single account. If we treat $\Theta$ and $\Phi$ as random variables, our target is to estimate the Maximum A Posteriori for the generative model.\par

@@

@@

Equation 2 represents the probability of assigning topic t and passenger p for i-th word in order o. $C^{WT}$ is the word-topic matrix, and $C_{wt}$ is the number of times word $w$ is assigned to topic $t$ except for the current word. $C^{TP}$ represents the topic-passenger matrix, and $C_{tp}$ indicates the number of times passenger $p$ is assigned to topic $t$ except for the current word $w_{oi}$. $W$ is the size of vocabulary, $T$ represents the number of topics and $P$ is the number of passengers. From the sampling metric, we can estimate the topic-word distribution and passenger-topic distribution:\\

@@

@@

@@

The algorithm performs in three steps, respectively, initialization, sampling and updating. At the first step, we assign each word in corpus with random passengers and topics. Then for every topic $t$, we count the frequency of each vocabulary item assigned with $t$. For every passenger $p$, we count the frequency of each $K$ topic assigned with $p$. For every sampling operation, a word in corpus is chosen, the probability of topic distribution and passenger distribution is calculated conditioned on the rest words in corpus applying Equation 2, with these two probability, we can sample a new topic and passenger for the current word. After several batch iterations, the passenger-topic matrix $\Theta$ and topic-word matrix $\Phi$ can be updated applying Equation 3,4. So the computational efficiency is the number of words multiplies number of topics, passengers, and iteration times.\par

@@

@@

In conclusion, the passenger prediction task can be partitioned into two steps. For the first step, a set of decisive factors are extracted by domain field knowledge to generate a vocabulary. And the parameter $\Theta$ and $\Phi$ are trained through several batches of Gibbs sampling. In the second step, passengers of an anonymous order can be predicted applying Equation 5. Algorithm 1 describes the predicting process mathematically.\par

>>>>

\documentclass{llncs}
\bibliographystyle{splncs}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=blue,citecolor=blue]{hyperref}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{document}

@@

%

\author{Yafeng Zhao, Jian Cao}
\institute{Department of Computer Science and Engineering,\\
Shanghai Jiao Tong University, Shanghai 200240, China\\
\email{zyfgs2012@163.com, cao-jian@sjtu.edu.cn}}

\maketitle
\begin{abstract}
Personalized recommendation is needed for online flight ticket booking service because it is a difficult task for a traveller to select the flight when the number of available flights is large. Traditionally, we can recommend flights for a user based on his historical orders from his account. However, people often book tickets for companions through his account. In this case, the preferences of other travellers should also be considered. Unfortuanately, before placing the order, people will not provider passengers' information. Therefore, we propose a probabilistic method for passenger prediction based on historical behaviors and contextual knowledge. We then experimentally demonstrate its effectiveness on a real dataset. The result shows that our method outperforms conventional methods.\\
\keywords{recommender system, probabilistic topic model, shared account}
\end{abstract}

\section{Introduction}
\label{sec:intro}
Recent years, with the rapid development of online travel agency, there are more and more passengers booking flight tickets through online service providers. Typically, an user may input his location, destination and departure time, then he will get a list of candidate flights. In general, there are dozens of items within a list, all of them are isomorphic. Thus, it takes quite long time for the user to compare and choose the most appropriate one. Therefore, a recommender system is quite necessary to be applied for better user experience. Figure \ref{fig:res} shows a list for all candidate flights departure from Beijing to Hong Kong when input a query. There are total 156 matchable flights in result. The website only provides simple sorting strategies such as departure time, arrival time, price. Thus, it may take quite long time for users to find the appropriate one.
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.48]{flight_tickets.eps}
\caption{Search result for flight tickets}
\label{fig:res}
\end{figure}\par
Essentially, an item recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that every account represents a single preference model. However, in online flight ticket booking, it is quite common that a user books tickets for the whole family or companions.  Obviously, every passenger has his own preferences. Unfortuanately, before placing the order, people will not provider passengers' information.  Therefore, if we want to improve the accuracy of recommendation, we have to predict who are going to fly.  \par
 +To predict who is using the account to place orders is regarded as shared account problem. Some approaches have been proposed to extract implicit user information within an account. For example,  in \cite{kabbur:nlmf}\cite{jason:embedding} nonlinear latent factorization is applied to  extract user information. However, these methods only rely on the analysis of relationships between products and users \cite{koen:top-n}\cite{yutaka:modeling}, and context information that can help differentiate users is neglected. Besides, current methods always treat an account as a fixed number of implicit users, and the experiment result can be only evaluated on artificial composite datasets. 
 +The flight booking service providers can acquire user's personal information only after he has submitted the order. Thus, our goal is to predict distribution of passengers in current session based on historical behavior and contextual knowledge so that we can make more pertinent recommendations. To achieve this target, we adopt the probabilistic author topic model to analyze passengers' behaviors. We regard all orders in an account as corpus, all passengers appearing in the account as authors, orders as documents, content attributes of the items and contextual knowledge as item list. Note that if an order has more than one passenger, they are co-authors of this document.\par
 +In experiment section, we apply our method to a real flight ticket order dataset.  And the result shows that our method has higher identification accuracy and a better recommendation accuracy.\par
 +This paper has following contributions:\\
\begin{enumerate}
+\item We provide a probabilistic passenger identification method for online flight booking service.
+\item We apply the passenger prediction results to flight recommendation.
+\item We evaluate our approach on a real flight order dataset, and the result shows that our approach can effectively improve the recommendation accuracy.
\end{enumerate}

\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in Section \ref{sec:relw}. In Section \ref{sec:mod}, we define the problem and describe the model. Besides, we also discuss how to do parameter inference. In section \ref{sec:exp}, we introduce the experiment dataset, evaluation metrics and the experimental results of passenge prediction and flight recommendation. Finally, we conclude the paper in section \ref{sec:con}.

\section{Related Work}
\label{sec:relw}
\textbf{Content based recommendation.} In online travel service industry such as airline tickets or hotel booking, the item information is usually highly structured, however, this information is quite mutable. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper representations of the attributes of items and the profile of user interests, then it tries to match  the profile of user's interests with attributes of items. In general, a content-based recommender system consists of three components. A content analyzer extracts structured attributes of items through a series of pre-processing steps. It produces the inputs for the system, which always need domain related knowledge. A profile learner analyzes and constructs user's profile of interests by collecting user's historical behaviors. In addition, this component also relies on feedbacks of users. A filtering component evaluates the relevance of items to particular users by matching them with users' profiles. The output for an item can be either binary or continuous, in this stage, some similarity measurements including cosine or euclidean distance can usually be applied.\par
\textbf{Flight Ticket Recommendation.} There is little work focusing on personalized flight ticket recommendation. Many online flight ticket booking agencies provide sorting, filtering strategies to improve user experience, such as sorting by price or takeoff time and filtering by airline or departure airport. Others also provide additional information that can help users, for example, placing the hottest flights at top or highlighting some of them. These methods do make sense in helping users find appropriate flights within less time. However, they can not provide personlized recommendations. In 2004, Lorcan Coyle\cite{lor:flight} proposed a Personal Travel Assistant(PTA). PTA is a method of case based reasoning, which takes user's historical orders as cases and recommends the most similar candidate flight to user's query. In addition, PTA also proposes attribute weight training and case storage method.\par
\textbf{Implicit feedback.} The user's feedback can be divided into implict  or explict type. The explicit feedback can be acquired from users' ratings or text comments. However,  users may not willing to provide explict feedback since it needs additional efforts. Moreover, in many applications, a single numeric rating may not be adequate for describing users feeling to an item. Implicit feedback methods\cite{song:mining} try to infer users' attitudes to an item based on his behaviors, such as clicking ,collecting or bookmarking et al. It has the advantage of being free from the problem of lacking feedback data. Nevertheless, implicit feedback based recommender suffers from problems of lack of negative feedbacks and data noises. In order to provide recommendations when only implict feedbacks are available, some methods are proposed, such as one class collaborative filtering and content based recommender.\par

\textbf{Probabilistic topic models.}  Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data, which are widely used in information retrieval, document annotation, etc. The main step is extracting latent factors from corpus, named topics, which are commonly probabilistic distribution over words. Topic models have three major advantages over other document modeling methods. At first, topic models provide a completely unsupervised approach to extract topics, thus requiring no document labels and no initialization. Unsupervised methods are necessary for modeling large document corpus, besides, in some fields, with the rapid change of content's topics, predefined topic categories may not reflect the development of specific field. Secondly, latent topics are interpretable respectively, so that extracted results are understandable for users. Finally, each document may consist of multiple topics, document can be considered as probabilistic distribution over topics. Many existing models regard documents as mixtures of topics, such as pLSI, LDA. The author topic model is a generative model that extends LDA to include authorship information for document modeling. In our approach, topic model is innovatively applied to indentify users. \par

\textbf{Shared account recommendation.} Some work aims at addressing the challenge of identifying current users who share a single account. A top-N recommendation for shared accounts approach was proposed by Koen et al.\cite{koen:top-n} It is an item-based top-N collaborative filtering recommender system on binary, positive-only feedback. Another method proposed by Santosh et al.\cite{kabbur:nlmf} models users within an account with a much richer representation. It uses a nonlinear matrix factorization methods for predicting the recommendation score. Yutaka et al.\cite{yutaka:modeling} introduced an approach for modeling multiple users' purchase in a single account using an extended pLSA model. All above methods are applied in scenario that users' personal information can't be acquired explicitly.

\section{User Prediction Model}
\label{sec:mod}
In this Section, we propose a generic model for predicting the probabilistic distribution of passengers under certain contextual environment. Finally we integrate user prediction into recommendation to get higher accuracy. Figure 2 shows an overview of the recommendation process. At first,a dataset for orders are stored, each account may contains several passengers. We use this data to train user prediction model as well as extract preference at user granularity. Before a recommendation process starts, we calculate user distribution probability applying trained model and contextual knowledge of current session. At the final step, we generate a ranked candidate list as the recommendation result based on user prediction result and user's preference.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of flight recommendation based on passenger prediction}
\label{fig:over}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items, it represents the latent feature. And an account is a multinomial distribution over topics, it can represent the preference of an account. Each purchase can be regared a sample that selects a topic z from account a, and takes an item from topic z.\par
When topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite different to be modeled. Conventional items only have static features, means that the content of an item is less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the ticket of different price as one item. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly.\par
Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. Some factors are discrete while others are continuous. For continuous factors, we can partition them into discrete non-overlapping intervals applying . Thus, the number of discrete alternatives of each factor is also very limited. We take orders within each account as a corpus. Our notations are summarized in table \ref{tab:not}.\par
\begin{table}[!htbp]
\centering
\caption{\label{tab:not}Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected Factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

@@

\begin{enumerate}
\item For each passenger $p=1,\dots,|P|$ draw $\Theta_p \sim Dirichlet(\alpha)$
\item For each topic $t=1,\dots,K$ draw $\Phi_t \sim Dirichlet(\beta)$
\item For each order $o=1,\dots,O$
       \begin{enumerate}[fullwidth,itemindent=1em,label=(\alph*)]
       \item given passengers $P$
       \item For each word $i=1,\dots, N_o$
              \begin{enumerate}[fullwidth,itemindent=2em,label=(\roman*)]
              \item draw a passenger $X_{oi} \sim Uniform(P)$
              \item draw a topic $Z_{oi} \sim Discrete(\theta_{X_{oi}})$
              \item draw a word $w_{oi} \sim Discrete(\phi_{Z_{oi}})$
              \end{enumerate}
       \end{enumerate}
\end{enumerate}
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.55]{f1.eps}
\caption{Probability Graph for the author-topic model}
\label{fig:mod}
\end{figure}\par
Figure \ref{fig:mod} shows the graphical model corresponding to this process, where shaded and non-shaded nodes represent observed and unobserved variables respectively. It's a plate notation, the number at the bottom of each box indicates repeated times of operation. Arrows represents the conditional dependency between variables. In the author-topic model, observed variables include words in an order and the passengers of the order.

\subsection{Parameter inference}
As mentioned above, the author-topic model includes two sets of unknown parameters, the $P$ passenger-topic distributions $\theta$, and the $K$ topic-word distributions $\phi$. There are a variety of algorithms that can be applied to estimate the parameters of author-topic models, the Expectation-Maximization algorithm and Gibbs sampling\cite{gregor:esti}. Generic EM algorithms are likely to face local maxima and computational inefficient. In this paper, we utilize Gibbs sampling, it does not explicitly estimate parameters, instead, it evaluates posteriori distribution just on drawn passenger $X$ and topic $Z$. Thus it is simple for Dirichlet priors.\par
We can obtain the probability of every word $\mathbf{w}_o$ generated in each order, conditioned on $\Theta$ and $\Phi$ is:\\
\begin{flalign}
\begin{split}
\label{eq:1} 
& P(\mathbf{w}_o | \Theta,\Phi,P) =\prod_{i=1}^{N_o}P(w_{oi}|\Theta,\Phi,\mathbf{p}_o) \\
& =\prod_{i=1}^{N_o}\sum_{p=1}^{|P|}\sum_{t=1}^{K}P(w_{oi}|z_{oi}=t,\Phi)
P(z_{oi}=t|x_{oi}=p,\Theta)P(x_{oi}=p|\mathbf{p}_o)\\
& =\prod_{i=1}^{N_o}\frac{1}{|P|}\sum_{p \in p_o}\sum_{t=1}^{K}\phi_{w_{oi}t}\theta_{tp}
\end{split} &
\end{flalign}

With the above generative model, $P(x_{oi}=p|\mathbf{p}_o)$ is assumed to be a uniform distribution over passenger list $P$. Each topic is drawn independently when conditioned on $\Theta$ and $p_o$, and each word is drawn independently when conditioned on $\Phi$ and $z$. Equation \ref{eq:1} can be applied as the likelihood of all orders in a single account. If we treat $\Theta$ and $\Phi$ as random variables, our target is to estimate the Maximum A Posteriori for the generative model.\par

@@

\label{eq:2}

@@

Equation \ref{eq:2} represents the probability of assigning topic t and passenger p for i-th word in order o. $C^{WT}$ is the word-topic matrix, and $C_{wt}$ is the number of times word $w$ is assigned to topic $t$ except for the current word. $C^{TP}$ represents the topic-passenger matrix, and $C_{tp}$ indicates the number of times passenger $p$ is assigned to topic $t$ except for the current word $w_{oi}$. $W$ is the size of vocabulary, $T$ represents the number of topics and $P$ is the number of passengers. From the sampling metric, we can estimate the topic-word distribution and passenger-topic distribution:\\

@@

\label{eq:3}

@@

\label{eq:4}

@@

The algorithm performs in three steps, respectively, initialization, sampling and updating. At the first step, we assign each word in corpus with random passengers and topics. Then for every topic $t$, we count the frequency of each vocabulary item assigned with $t$. For every passenger $p$, we count the frequency of each $K$ topic assigned with $p$. For every sampling operation, a word in corpus is chosen, the probability of topic distribution and passenger distribution is calculated conditioned on the rest words in corpus applying Equation \ref{eq:2}, with these two probability, we can sample a new topic and passenger for the current word. After several batch iterations, the passenger-topic matrix $\Theta$ and topic-word matrix $\Phi$ can be updated applying Equation \ref{eq:3}, \ref{eq:4}. So the computational efficiency is the number of words multiplies number of topics, passengers, and iteration times.\par

@@

\label{eq:5}

@@

In conclusion, the passenger prediction task can be partitioned into two steps. For the first step, a set of decisive factors are extracted by domain field knowledge to generate a vocabulary. And the parameter $\Theta$ and $\Phi$ are trained through several batches of Gibbs sampling. In the second step, passengers of an anonymous order can be predicted applying Equation \ref{eq:5}. Algorithm \ref{alg:1} describes the predicting process mathematically.\par
